Rewind: Automatically Generated\\Digital Memories from Mobile Geolocation Data

### 2016-09-21 paper/uist_2016/paper.tex
memorability 1/6

\section{Memorability}
Rewinds, and digitally-mediated memories in general, are meant to engage a person's memory. To find out what features make a digital memory memorable, we examined participant responses to whether or not they were able to remember the location shown in a photo (semantic memory) or recall the trip portrayed by a Rewind (episodic memory). Photo memorability is determined by whether or not the participant remembered the location depicted in a photo. Because a Rewind contains many locations, it would be unfair to consider the memorability of Rewind based on whether the participant remembers any locations in the Rewind. So, Rewind memorability is determined by whether or not the participant remembered and could distinguish the particular trip portrayed in a Rewind. In this section, we will look at differences in photo and Rewind memorability based on the five metrics discussed.

\subsubsection{Semantic Memory} 
Within-subject comparison in our study reveals that participants are significantly better at recalling semantic memory of a particular location with \textit{Rewind} videos than with static street-view images, $\chi^2(1, N=356) = 96.987$, $p <.01$. When being presented with a street-side image retrieved based on their location data, participants were able to recognize places visited 50\% of the time; yet recognition rate of locations reached 97\% when participants started playing accompanied Rewinds, with a significant increase of 47\%. It is perhaps not too surprising as cinematic media like Rewinds would provide users with more contextual meanings and retrival cues for semantic processing, as three participants (P4, P6 and P10) have once expressed their contentment after watching Rewinds by claiming ``I recognize the photo now!'' when their initial attempts with static images recognition were fruitless.

%
\subsubsection{Episodic Memory}
It could possibly be inferred that participants are likely to benefit more from Rewinds than static street-view images at recalling episodic memory. Although participants were not explicitly asked about trip rememberance when presented with static images as opposed to what they were asked after watchiing Rewinds, location rememberance from static images may serve as an upper-bound estimation of how much an individual could recall about a particular trip; since if participants fail to form any cognitive association with the locations depicted in the images, it is very unlikely that they would be able to associate target itineraries with the locations shown. Thus, it is perhaps safe to presume that location rememberance with images bounds above the maximum likelihood of trip rememberance with the same images presented to the participants, and we used location rememberance from image recognition as an upper-limit gauge against trip rememberance from Rewinds thereafter.

% h - add another table, only abt memorability tho
% explain distance groups in the beginning
% where is home? need to explain this part in system?
Using the aforementioned presumption, our study examined effects of Rewinds videos on episodic processing in metric-specific groups, as well as the performance overall. Our study shows that comparing to static street-view images, recollection of event-driven memories using \textit{Rewind} videos is significantly improved for trips taken that were 100 - 1,000 km (62 - 620 miles) away from home by 50\%, $\chi^2(1, N=20) = 3.215$, $p <.05$; and for trips that were least frequently taken by about 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$. Our study also suggested that GPS users would benefit more from Rewinds, even though the difference was mildly significant by an increase about 15\%, $\chi^2(1, N=118) = 2.145$, $p = .0715 (<.1)$. 
% h - prob explain this a little bit more

Overall, trip rememberance from Rewinds is about 4\% higher than that from static street-view images. Although chi-squared one-tailed test did not reveal this difference to be statistically significant, $\chi^2(1, N=356) = 0.41$, $p >.1$, we are positive that the actual difference would be more significant since the trip rememberance from images was a best possible estimation from location rememberance.

\subsubsection{Believability and Desirability}
Although not statistically significant, %TODO TODO TODO TODO TODO

% shall we change photo/rewind video rememberance to location/trip rememberance? april - i define photo/rewind remembrance below. a photo and a rewind can have both location and trip remembrance. for instance from a photo you might remember the location but not the trip. 
\subsection{Memorability}

%TODO heather - didn't do the stats for video location vs photo locations, not sure if it is necessary? - photo vs video stats?
%TODO will add this part back in within-group comparison

% \subsubsection{Rewinds versus Photos}
% When participants were shown a street side photo from their location history, they were asked whether they recognized the location in the photo. In 50\% of cases, participants said they did. When they were shown the Rewind associated with that photo, nearly every case had recognizable locations---97\%. This result is expected since Rewinds contain more location information than static photos, so if participants do not recognize some depicted locations, they still have many other chances to pick out locations they remember. Rewind memorability as calculated by whether participants could identify the particular trip they were on from watching the Rewind was 54\%. However, there were cases where the participant would only remember the photo after seeing its accompanying Rewind, which provided a context. More than one participant exclaimed after watching a Rewind, ``I recognize the photo now!''

\subsubsection{Distance}
% only affect location rememberance
% 0 - 100%, 1 - 75%, 2 - 72%, 3 - 44%, 4 - 52%, 5 - 10%, $\chi^2(5,179) = 18.106$, p < 0.01
The distance away from the participant’s home had differing effects on photo memorability and Rewind memorability (Figure~\ref{fig:distancememory}).

@@

\subsubsection{GPS Usage}
% Our findings show that GPS usage significantly affect the rememberance of locations $\chi^2(1,179) = 8.734$, p < 0.01, but does not impose a siginificant effect on trip rememberance. (GPS users remember locations 33% of the time whereas non-GPS users remember locations 58% of the time)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% heather - come back to this if time allows
% because memories are simply not encoded so there's no way to retrieve it in the first place?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Our findings show that GPS usage significantly affect participants' remembrance of locations $\chi^2(1,179) = 8.734$, p = 0.0016 ($< 0.01$), but does not impose significant effects on trip remembrance $\chi^2(1,177) = 0.713$, p = 0.1992 ($> 0.05$).  When asked whether they remember the location in a static photo, non-regular GPS users gave a positive answer 58\% of the time whereas regular GPS users gave a positive answer 33\% of the time. Although the difference between non-regular and regular GPS users' trip remembrance is not statistically significance, the between-subject percentage values still lean towards the observation that it is perhaps easier for non-GPS users to recall episodic memories from external memory cues, given that the total trip remembrance rate for non-GPS users is 48\% and that of GPS users is 56\%. Since regular GPS users are frequently directed step-by-step on how to navigate through their environment, regular GPS users may not need to pay as much attention to their surroundings, leading to potentially diminished locational awareness which could be the reason why regular GPS users had a lower remembrance rate for their ambience, particularly locations. \\

@@

>>>>

\section{Memorability}
Rewinds, and digitally-mediated memories in general, are meant to engage a person's memory. To find out what features make a digital memory memorable, we examined participant responses to whether or not they were able to remember the location shown in a photo (semantic memory) or recall the trip portrayed by a Rewind (episodic memory). Photo memorability is determined by whether or not the participant remembered the location depicted in a photo. Because a Rewind contains many locations, it would be unfair to evaluate the memorability of Rewind based on whether the participant remembers any locations in the Rewind. If participants do not recognize some depicted locations, they still have many other chances to pick out locations they remember. So, Rewind memorability is determined by whether or not the participant remembered and could distinguish the particular trip portrayed in a Rewind. In this section, we will look at differences in photo and Rewind memorability based on the five metrics discussed previously.

\subsubsection{Rewinds vs. Static Photos} 
Our study reveals that participants have higher semantic memory of locations when presented with Rewinds than with static street-view photos ($\chi^2(1, N=356) = 96.987$, $p <.01$). When presented with a street-view photo generated from the participant's location data, participants were able to recognize the depicted location 49.7\% of the time. Recognition rate of locations reached 96.6\% when participants were shown the accompanying Rewind. As discussed above, this can be expected because cinematic media like Rewinds provide more context and retrieval cues. Three participants (P4, P6, and P10) who were unable to recall a location from a static photo actually recognized the location after seeing the accompanying Rewind. They expressed their contentment after watching the Rewind, proclaiming ``I recognize the photo now!'' When we evaluate Rewind memorability based on episodic memory rather than semantic memory, the results show that participants recalled the exact trip depicted by the Rewind 53.7\% of the time.
%TODO make sure episodic memory is discussed in discussion section--visit frequency diminishes episodic memory

%TODO h - add another table, only abt memorability tho
%TODO  where is home? need to explain this part in system? make sure this is explained
\subsubsection{Distance}
% only affect location rememberance
% 0 - 100%, 1 - 75%, 2 - 72%, 3 - 44%, 4 - 52%, 5 - 10%, $\chi^2(5,179) = 18.106$, p < 0.01
The distance away from the participant’s home had differing effects on photo memorability and Rewind memorability (Figure~\ref{fig:distancememory}).

@@

Our study examined episodic processing in metric-specific groups, as well as the performance overall. Our study shows that comparing to static street-view images, recollection of event-driven memories using \textit{Rewind} videos is significantly improved for trips taken that were 100 - 1,000 km (62 - 620 miles) away from home by 50\%, $\chi^2(1, N=20) = 3.215$, $p <.05$; and for trips that were least frequently taken by about 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$.

\subsubsection{GPS Usage}
Our study also suggested that GPS users would benefit more from Rewinds, even though the difference was mildly significant by an increase about 15\%, $\chi^2(1, N=118) = 2.145$, $p = .0715 (<.1)$. 
% h - prob explain this a little bit more

% Our findings show that GPS usage significantly affect the rememberance of locations $\chi^2(1,179) = 8.734$, p < 0.01, but does not impose a siginificant effect on trip rememberance. (GPS users remember locations 33% of the time whereas non-GPS users remember locations 58% of the time)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% heather - come back to this if time allows
% because memories are simply not encoded so there's no way to retrieve it in the first place?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Our findings show that GPS usage significantly affect participants' remembrance of locations $\chi^2(1,179) = 8.734$, p = 0.0016 ($< 0.01$), but does not impose significant effects on trip remembrance $\chi^2(1,177) = 0.713$, p = 0.1992 ($> 0.05$).  When asked whether they remember the location in a static photo, non-regular GPS users gave a positive answer 58\% of the time whereas regular GPS users gave a positive answer 33\% of the time. Although the difference between non-regular and regular GPS users' trip remembrance is not statistically significance, the between-subject percentage values still lean towards the observation that it is perhaps easier for non-GPS users to recall episodic memories from external memory cues, given that the total trip remembrance rate for non-GPS users is 48\% and that of GPS users is 56\%. Since regular GPS users are frequently directed step-by-step on how to navigate through their environment, regular GPS users may not need to pay as much attention to their surroundings, leading to potentially diminished locational awareness which could be the reason why regular GPS users had a lower remembrance rate for their ambience, particularly locations. \\

@@

%TODO where to put this????  maybe split discussion into three sections on mem/believ/desir
\subsubsection{Episodic Memory}
It could possibly be inferred that participants are likely to benefit more from Rewinds than static street-view images at recalling episodic memory. Although participants were not explicitly asked about trip rememberance when presented with static images as opposed to what they were asked after watchiing Rewinds, location rememberance from static images may serve as an upper-bound estimation of how much an individual could recall about a particular trip; since if participants fail to form any cognitive association with the locations depicted in the images, it is very unlikely that they would be able to associate target itineraries with the locations shown. Thus, it is perhaps safe to presume that location rememberance with images bounds above the maximum likelihood of trip rememberance with the same images presented to the participants, and we used location rememberance from image recognition as an upper-limit gauge against trip rememberance from Rewinds thereafter.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% end where to put this???
